Submission 1.1

### 2023-08-19 Submission 1.1.tex
Homework 2 - Knights and Knaves - Fixed validity on some problems (contradictory -> valid but not sound) - Truth tables (part d)

\documentclass{article}
\usepackage{enumerate}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage[table]{xcolor}
\title{Submission 1.1}
\author{Saffron Liu}
\date{}
\begin{document}
\maketitle
\begin{flushleft}
    For problems 2-12:\\
\end{flushleft}
\begin{enumerate}[(a)]
    \item \textit{Write the argument in standard form.}
    \item \textit{Claim whether the argument is valid or sound. In some cases, soundness will be difficult to determine, so “soundness is difficult to determine” is an appropriate answer.}
    \item \textit{Translate the argument into truth-functional logic. Be sure to delineate the extensions you give to atomic sentences and write the argument (now in truth-functional logic) in standard form.}
          % \item \textit{Prove that the argument is valid or invalid (as appropriate) with a proof by truth table.}
          % \item \textit{Prove that the argument is valid or invalid (as appropriate) with an informal proof.}
\end{enumerate}
\begin{flushleft}
    Problems:
\end{flushleft}
\begin{enumerate}
    \setcounter{enumi}{1}
    \item I have already said that he must have gone to King's Pyland or to Mapleton. He is not at King's Pyland, therefore he is at Mapleton. (Sir Arthur Conan Doyle) (B 20)
          \begin{enumerate}[(a)]
              \item He went to King's Pyland or to Mapleton.\\
                    He is not at King's Pyland.\\
                    \rule{15em}{.5pt}\\
                    He is at Mapleton.
              \item The argument not valid, because he might have gone to King's Pyland or Mapleton, but no longer be in either of those places. Because it is not valid, it cannot be sound.
              \item \textit{A}: He went to King's Pyland\\
                    \textit{B}: He went to Mapleton\\
                    \textit{C}: He is at King's Pyland\\
                    \textit{D}: He is at Mapleton\\\\
                    $A \lor B$\\
                    $\neg C$\\
                    \rule{5em}{.5pt}\\
                    $D$
          \end{enumerate}
    \item The patient will die unless we operate. We will operate. Therefore the patient will not die. (B 20)
          \begin{enumerate}[(a)]
              \item The patient will die unless we operate.\\
                    We will operate.\\
                    \rule{15em}{.5pt}\\
                    The patient will not die.
              \item The argument is not valid, because while operating is necessary for the patient to survive, it may not be sufficient. Since the argument is not valid, it cannot be sound.
              \item \textit{A}: We operate\\
                    \textit{B}: The patient dies\\

                    $\neg A \to B$\\
                    $A$\\
                    \rule{5em}{.5pt}\\
                    $\neg B$
          \end{enumerate}
    \item If I'm right, then I'm a fool. But if I'm a fool, I'm not right. Therefore, I'm no fool. (B 70)
          \begin{enumerate}[(a)]
              \item If I'm right, then I'm a fool.\\
                    If I'm a fool, then I'm not right.\\
                    \rule{15em}{.5pt}\\
                    I'm not a fool.
              \item The argument is neither valid nor sound because the premises form a contradiction. Being right implies being a fool, but being a fool implies being not right, so being right implies being not right.
              \item \textit{A}: I'm right\\
                    \textit{B}: I'm a fool\\

                    $A \to B$\\
                    $B \to \neg A$\\
                    \rule{5em}{.5pt}\\
                    $\neg B$
          \end{enumerate}
    \item If I'm right, then I'm a fool. But if I'm a fool, I'm not right. Therefore, I'm not right. (B 70)
          \begin{enumerate}[(a)]
              \item If I'm right, then I'm a fool.\\
                    If I'm a fool, then I'm not right.\\
                    \rule{15em}{.5pt}\\
                    I'm not right.
              \item Again, the premises form a contradiction and so the argument can't be either valid or sound.\\  
              \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory? On one hand, if you're right you're a fool and if you're a fool you're not right doesn't seem to work, but it is possible to be a fool and not right without violating either premise. So it is possible for them to both be true at the same time, I think? Just very strange.\\
              Update: Contradictory means you can prove $A \to \neg A$, regardless of whether or not there is a case where all premises are true.}
              \item \textit{A}: I'm right\\
                    \textit{B}: I'm a fool\\

                    $A \to B$\\
                    $B \to \neg A$\\
                    \rule{5em}{.5pt}\\
                    $\neg A$
          \end{enumerate}
    \item If Einstein's theory of relativity is correct, light bends in the vicinity of the sun. Light does indeed bend at the vicinity of the sun. It follows that Einstein's theory of relativity is correct. (B 70)
          \begin{enumerate}[(a)]
              \item If Einstein's theory of relativity is correct, then light bends in the vicinity of the sun.\\
                    Light bends in the vicinity of the sun.\\
                    \rule{15em}{.5pt}\\
                    Einstein's theory of relativity is correct.
              \item The argument is not valid because light could bend near the sun for a different reason than Einstein posits, so the premises would be true but the conclusion would be false. Because it is not valid, it is also not sound.
              \item \textit{A}: Einstein's theory of relativity is correct\\
                    \textit{B}: Light bends in the vicinity of the sun

@@

                    $\neg B \to \neg A$\\
                    $\neg B$\\
                    \rule{5em}{.5pt}\\
                    $\neg A$
          \end{enumerate}
    \item If you are ambitious, you'll never achieve all your goals. But life has meaning only if you have ambition. Thus, if you achieve all your goals, life has no meaning. (B 132)
          \begin{enumerate}[(a)]
              \item If you are ambitious, you'll never achieve all your goals.\\
                    Life has meaning only if you have ambition.\\
                    \rule{15em}{.5pt}\\
                    If you achieve all your goals, then life has no meaning.
              \item The argument is valid. Soudness is difficult to determine.
              \item \textit{A}: You are ambitious.\\
                    \textit{B}: You achieve all your goals.\\
                    \textit{C}: Life has meaning.\\

@@

                    $A \to B$\\
                    $C \to A$\\
                    \rule{5em}{.5pt}\\
                    $C \to B$
          \end{enumerate}
    \item "Thin is guilty," observed Watson, “because either Holmes is right and the vile Moriarty is guilty or he is wrong and the scurrilous Thin did the job; but those scoundrels are either both guilty or both innocent; and, as usual, Holmes is right.” (J 18)
          \begin{enumerate}[(a)]
              \item Holmes is right or wrong.\\
                    If Holmes is right, Moriarty is guilty.\\
                    If Holmes is wrong, Thin is guilty.\\
                    Moriarty and Thin are either both guilty or both innocent (i.e. If Moriarty is guilty, Thin is guilty).\\
                    \rule{15em}{.5pt}\\
                    Thin is guilty.
              \item The argument is valid. Soundness is difficult to determine.
              \item \textit{A}: Holmes is right.\\
                    \textit{B}: Moriarty is guilty.\\
                    \textit{C}: Thin is guilty.\\

@@

                    $A \iff B$\\
                    $A \land \neg B$\\
                    \rule{5em}{.5pt}\\
                    $C$
          \end{enumerate}
    \item God is omnipotent if and only if He can do everything. If He can't make a stone so heavy that He can't lift it, then he can't do everything. But if He can make a stone so heavy that He can't lift it, He can't do everything. Therefore, either God is not omnipotent or God does not exist. (B 132)
          \begin{enumerate}[(a)]
              \item God is omnipotent if and only if he can do everything.\\
                    If he cannot make a stone so heavy that he cannot lift it, he cannot do everything.\\
                    If he can make a stone that he cannot lift, he cannot do everything.\\
                    \rule{15em}{.5pt}\\
                    Either god is not omnipotent or he does not exist.
              \item The argument is valid. Soudness is difficult to determine.\\
                    \textit{Saffron's Commentary: I have no idea where the existence of God fits into that argument. In the English, it makes sense that it suddenly appears in the conclusion; in the logical notation, less so. I don't know whether that makes it invalid or not.\\
                    Update: It seems like there's a contradiction posed by $\neg D \to \neg C$ and $D \to \neg C$, because the second statemnent is equivalent to $C \to \neg D$, meaning $C \to \neg C$. However, Descartes'} cogito \textit{has the same issue.}
              \item \textit{A}: God exists.\\
                    \textit{B}: God is omnipotent.\\
                    \textit{C}: God can do everything.\\
                    \textit{D}: God can make a stone which he cannot lift.\\

>>>>

\documentclass{article}
\usepackage{enumerate}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage[table]{xcolor}

\title{Submission 1.1}
\author{Saffron Liu}
\date{}
\begin{document}
\maketitle

\begin{enumerate}
      \item Knaves always lie, knights always tell the truth, and in Transylvania, where everybody is one or the other (but you can’t tell which by looking), you encounter two people, one of whom says, “He’s a knight or I’m a knave.” What are they? (J 20)\\\\
            A: I'm a knight.\\
            B: He's a knight.\\
            C: He's a knight or I'm a knave.\\
            Notes:\\
            If being a knight is false, being a knave is true.\\
            If I am a knight, then (he's a knight or I'm not a knight). If I'm not a knight, then not (he's a knight or I'm not a knight).\\
            To solve the problem, we must find a combination of knight and knave such that the truth value of "he's a knight or I'm a knave" is consistent with whether or not I'm a knight. Therefore, we're searching for a value where C and A are equivalent.\\
            \begin{tabular}{c|c|c|c}
                  A & B & $C: B \lor \neg A$ & $A \leftrightarrow C$ \\
                  \hline
                  T & T & T                  & T                     \\
                  T & F & F                  & F                     \\
                  F & T & T                  & F                     \\
                  F & F & T                  & F                     \\
            \end{tabular}
            Both he and I are knights.
\end{enumerate}
\begin{flushleft}
      For problems 2-12:\\
\end{flushleft}
\begin{enumerate}[(a)]
      \item \textit{Write the argument in standard form.}
      \item \textit{Claim whether the argument is valid or sound. In some cases, soundness will be difficult to determine, so “soundness is difficult to determine” is an appropriate answer.}
      \item \textit{Translate the argument into truth-functional logic. Be sure to delineate the extensions you give to atomic sentences and write the argument (now in truth-functional logic) in standard form.}
      \item \textit{Prove that the argument is valid or invalid (as appropriate) with a proof by truth table.}
            % \item \textit{Prove that the argument is valid or invalid (as appropriate) with an informal proof.}
\end{enumerate}
\begin{enumerate}
      \setcounter{enumi}{1}
      \item I have already said that he must have gone to King's Pyland or to Mapleton. He is not at King's Pyland, therefore he is at Mapleton. (Sir Arthur Conan Doyle) (B 20)
            \begin{enumerate}[(a)]
                  \item He went to King's Pyland or to Mapleton.\\
                        He is not at King's Pyland.\\
                        \rule{15em}{.5pt}\\
                        He is at Mapleton.
                  \item The argument not valid, because he might have gone to King's Pyland or Mapleton, but no longer be in either of those places. Because it is not valid, it cannot be sound.
                  \item \textit{A}: He went to King's Pyland\\
                        \textit{B}: He went to Mapleton\\
                        \textit{C}: He is at King's Pyland\\
                        \textit{D}: He is at Mapleton\\\\
                        $A \lor B$\\
                        $\neg C$\\
                        \rule{5em}{.5pt}\\
                        $D$
            \end{enumerate}
      \item The patient will die unless we operate. We will operate. Therefore the patient will not die. (B 20)
            \begin{enumerate}[(a)]
                  \item The patient will die unless we operate.\\
                        We will operate.\\
                        \rule{15em}{.5pt}\\
                        The patient will not die.
                  \item The argument is not valid, because while operating is necessary for the patient to survive, it may not be sufficient. Since the argument is not valid, it cannot be sound.
                  \item \textit{A}: We operate\\
                        \textit{B}: The patient dies\\

                        $\neg A \to B$\\
                        $A$\\
                        \rule{5em}{.5pt}\\
                        $\neg B$
                  \item
                        \begin{tabular}{c|c|c}
                              A                      & B & $\neg A \to B$ \\
                              \hline
                              T                      & T & T              \\
                              T                      & F & T              \\
                              \rowcolor{lightgray} F & T & T              \\
                              \rowcolor{lightgray} F & F & F
                        \end{tabular} \\
                        Where both premises are true, B can still be either true or false. This means there is no guarantee the patient will not die, so the argument is invalid.
            \end{enumerate}
      \item If I'm right, then I'm a fool. But if I'm a fool, I'm not right. Therefore, I'm no fool. (B 70)
            \begin{enumerate}[(a)]
                  \item If I'm right, then I'm a fool.\\
                        If I'm a fool, then I'm not right.\\
                        \rule{15em}{.5pt}\\
                        I'm not a fool.
                  \item The argument is valid but not sound because the premises form a contradiction. Being right implies being a fool, but being a fool implies being not right, so being right implies being not right.
                  \item \textit{A}: I'm right\\
                        \textit{B}: I'm a fool\\

                        $A \to B$\\
                        $B \to \neg A$\\
                        \rule{5em}{.5pt}\\
                        $\neg B$
            \end{enumerate}
      \item If I'm right, then I'm a fool. But if I'm a fool, I'm not right. Therefore, I'm not right. (B 70)
            \begin{enumerate}[(a)]
                  \item If I'm right, then I'm a fool.\\
                        If I'm a fool, then I'm not right.\\
                        \rule{15em}{.5pt}\\
                        I'm not right.
                  \item Again, the premises form a contradiction and so the argument is valid but not sound.\\
                        \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory? On one hand, if you're right you're a fool and if you're a fool you're not right doesn't seem to work, but it is possible to be a fool and not right without violating either premise. So it is possible for them to both be true at the same time, I think? Just very strange.\\
                              Update: Contradictory means you can prove $A \to \neg A$, regardless of whether or not there is a case where all premises are true.}
                  \item \textit{A}: I'm right\\
                        \textit{B}: I'm a fool\\

                        $A \to B$\\
                        $B \to \neg A$\\
                        \rule{5em}{.5pt}\\
                        $\neg A$
                  \item
                        \begin{tabular}{c|c|c|c}
                              A                      & B & $A \to B$ & $B \to \neg A$ \\
                              \hline
                              \rowcolor{lightgray} T & T & T         & F              \\
                              \rowcolor{lightgray} T & F & F         & T              \\
                              F                      & T & T         & T              \\
                              F                      & F & T         & T              \\
                        \end{tabular} \\
                        The argument is valid because in all rows where the premises are met, I am not right.\\
                        \textit{Saffron's Commentary: I'm not actually sure how to represent a contradiction in a truth table. It would work if there were no case where the contradictory statements could be true at the same time, but that's not the case here.}
            \end{enumerate}
      \item If Einstein's theory of relativity is correct, light bends in the vicinity of the sun. Light does indeed bend at the vicinity of the sun. It follows that Einstein's theory of relativity is correct. (B 70)
            \begin{enumerate}[(a)]
                  \item If Einstein's theory of relativity is correct, then light bends in the vicinity of the sun.\\
                        Light bends in the vicinity of the sun.\\
                        \rule{15em}{.5pt}\\
                        Einstein's theory of relativity is correct.
                  \item The argument is not valid because light could bend near the sun for a different reason than Einstein posits, so the premises would be true but the conclusion would be false. Because it is not valid, it is also not sound.
                  \item \textit{A}: Einstein's theory of relativity is correct\\
                        \textit{B}: Light bends in the vicinity of the sun

@@

                        $\neg B \to \neg A$\\
                        $\neg B$\\
                        \rule{5em}{.5pt}\\
                        $\neg A$
                  \item \begin{tabular}{c|c|c}
                              A                      & B & $(\neg B \to \neg A) = (A \to B)$ \\
                              \hline
                              \rowcolor{lightgray} T & T & T                                 \\
                              \rowcolor{lightgray} T & F & F                                 \\
                              \rowcolor{lightgray} F & T & T                                 \\
                              F                      & F & T
                        \end{tabular} \\ The argument is valid because the only row where both premises are true is where Congress does not agree to the cut.
            \end{enumerate}
      \item If you are ambitious, you'll never achieve all your goals. But life has meaning only if you have ambition. Thus, if you achieve all your goals, life has no meaning. (B 132)
            \begin{enumerate}[(a)]
                  \item If you are ambitious, you'll never achieve all your goals.\\
                        Life has meaning only if you have ambition.\\
                        \rule{15em}{.5pt}\\
                        If you achieve all your goals, then life has no meaning.
                  \item The argument is valid. Soudness is difficult to determine.
                  \item \textit{A}: You are ambitious.\\
                        \textit{B}: You achieve all your goals.\\
                        \textit{C}: Life has meaning.\\

@@

                        $A \to B$\\
                        $C \to A$\\
                        \rule{5em}{.5pt}\\
                        $C \to B$
                  \item
                        \begin{tabular}{c|c|c|c|c}
                              A                      & B & C & $A \to B$ & $C \to A$ \\
                              T                      & T & T & T         & T         \\
                              T                      & T & F & T         & T         \\
                              \rowcolor{lightgray} T & F & T & F         & T         \\
                              \rowcolor{lightgray} T & F & F & F         & T         \\
                              \rowcolor{lightgray} F & T & T & T         & F         \\
                              F                      & T & F & T         & T         \\
                              \rowcolor{lightgray} F & F & T & T         & F         \\
                              F                      & F & F & T         & T         \\
                        \end{tabular} \\ In the rows where both premises are met, there is only one case where Brown dies before the election, and in this case he does indeed retire to private life. This means the argument is valid.
            \end{enumerate}
      \item "Thin is guilty," observed Watson, “because either Holmes is right and the vile Moriarty is guilty or he is wrong and the scurrilous Thin did the job; but those scoundrels are either both guilty or both innocent; and, as usual, Holmes is right.” (J 18)
            \begin{enumerate}[(a)]
                  \item Holmes is right or wrong.\\
                        If Holmes is right, Moriarty is guilty.\\
                        If Holmes is wrong, Thin is guilty.\\
                        Moriarty and Thin are either both guilty or both innocent (i.e. If Moriarty is guilty, Thin is guilty).\\
                        \rule{15em}{.5pt}\\
                        Thin is guilty.
                  \item The argument is valid. Soundness is difficult to determine.
                  \item \textit{A}: Holmes is right.\\
                        \textit{B}: Moriarty is guilty.\\
                        \textit{C}: Thin is guilty.\\

@@

                        $A \iff B$\\
                        $A \land \neg B$\\
                        \rule{5em}{.5pt}\\
                        $C$
                  \item \begin{tabular}{c|c|c|c|c}
                              A & B & C & $A \iff B$ & $A \land \neg B$ \\
                              \hline
                              \rowcolor{lightgray} T & T & T & T          & F                \\
                              \rowcolor{lightgray} T & T & F & T          & F                \\
                              \rowcolor{lightgray} T & F & T & F          & T                \\
                              \rowcolor{lightgray} T & F & F & F          & T                \\
                              \rowcolor{lightgray} F & T & T & F          & F                \\
                              \rowcolor{lightgray} F & T & F & F          & F                \\
                              \rowcolor{lightgray} F & F & T & T          & F                \\
                              \rowcolor{lightgray} F & F & F & T          & F                \\
                        \end{tabular}\\
                        There is no case where all premises are true, so the premises are contradictory and thus the argument is valid.
            \end{enumerate}
      \item God is omnipotent if and only if He can do everything. If He can't make a stone so heavy that He can't lift it, then he can't do everything. But if He can make a stone so heavy that He can't lift it, He can't do everything. Therefore, either God is not omnipotent or God does not exist. (B 132)
            \begin{enumerate}[(a)]
                  \item God is omnipotent if and only if he can do everything.\\
                        If he cannot make a stone so heavy that he cannot lift it, he cannot do everything.\\
                        If he can make a stone that he cannot lift, he cannot do everything.\\
                        \rule{15em}{.5pt}\\
                        Either god is not omnipotent or he does not exist.
                  \item The argument is valid. Soudness is difficult to determine.\\
                        \textit{Saffron's Commentary: I have no idea where the existence of God fits into that argument. In the English, it makes sense that it suddenly appears in the conclusion; in the logical notation, less so. I don't know whether that makes it invalid or not.\\
                              Update: It seems like there's a contradiction posed by $\neg D \to \neg C$ and $D \to \neg C$, because the second statemnent is equivalent to $C \to \neg D$, meaning $C \to \neg C$. However, Descartes'} cogito \textit{has the same issue.}
                  \item \textit{A}: God exists.\\
                        \textit{B}: God is omnipotent.\\
                        \textit{C}: God can do everything.\\
                        \textit{D}: God can make a stone which he cannot lift.\\
